Finch: A Datastructure-Driven Array Programming Language

### 2024-03-20 sections/eval.tex
Update on Overleaf.

\subsubsection{Tensor Formats}
We benchmarked the SpMV kernel using the Dense level format for the outer level and the SparseList, SparseVBL, SparseBand, and SparsePoint Finch levels for the inner level of the sparse matrix. To represent scalar values, we benchmarked with both the Element and Pattern levels. The SparseList, SparseVBL, and SparseBand formats were functional for all matrices, but the SparsePoint format was applicable only for the synthetic permutation matrices. The Element format was applicable to represent scalars in all matrices, while the Pattern level was applicable only for boolean matrices. We tested with both the row-major and the column-major implementations (labeled with a "R" and "C", respectively, where applicable in Figure \ref{spmv_grouped}) with Finch.

We found that the SpMV performance was superior for the level format that best paralleled the structure of the tensor. Namely, matrices with a clear blocked structure like exdata\_1, TSOPF\_RS\_b678\_c1, and heart3 performed notably well with the SparseVBL format in comparison to both TACO and the Finch SparseList formats (which we consider the closest analog to the sparse matrix format used in TACO and other libraries) with speedups of 2.16, 1.55, and 1.30 relative to TACO and of 1.47, 2.13, and 1.42 relative to the fastest SparseList Finch SpMV implementation. Furthermore, the synthetic Toeplitz banded matrices we constructed performed the best with the SparseBand matrix, in particular with the $toeplitz\_large\_band$ and the $toeplitz\_medium\_band$ matrices having a speedup of 1.98 and 1.64 relative to TACO and a speedup of 1.47 and 1.24 relative to the fastest SparseList Finch SpMV implementation.

There were also significant advantages of using the Pattern format instead of the Element format to represent scalar values in the matrices when these values were boolean. The matrices in the SNAP collection which represent graph datasets are boolean and had an average speedup of X over TACO and X over the same Finch program, but with an Element format.

>>>>

\subsubsection{Tensor Formats}
We benchmarked the SpMV kernel using the Dense level format for the outer level and the SparseList, SparseVBL, SparseBand, and SparsePoint Finch levels for the inner level of the sparse matrix. To represent scalar values, we benchmarked with both the Element and Pattern levels. The SparseList, SparseVBL, and SparseBand formats were functional for all matrices, but the SparsePoint format was applicable only for the synthetic permutation matrices. The Element format was applicable to represent scalars in all matrices, while the Pattern level was applicable only for boolean matrices. We tested with both the row-major and the column-major implementations (labeled with a "R" and "C", respectively, where applicable in Figure \ref{spmv_grouped}) with Finch. Column-major SpMV consistently performs better than row-major SpMV (an average of 1.36x better) in TACO so we use column-major SpMV in TACO as our baseline.

We found that the SpMV performance was superior for the level format that best paralleled the structure of the tensor. Namely, matrices with a clear blocked structure like exdata\_1, TSOPF\_RS\_b678\_c1, and heart3 performed notably well with the SparseVBL format in comparison to both TACO and the Finch SparseList formats (which we consider the closest analog to the sparse matrix format used in TACO and other libraries) with speedups of 2.16, 1.55, and 1.30 relative to TACO and of 1.47, 2.13, and 1.42 relative to the fastest SparseList Finch SpMV implementation. Furthermore, the synthetic Toeplitz banded matrices we constructed performed the best with the SparseBand matrix, in particular with the toeplitz\_large\_band and the toeplitz\_medium\_band matrices having a speedup of 1.98 and 1.64 relative to TACO and a speedup of 1.47 and 1.24 relative to the fastest SparseList Finch SpMV implementation.

There were also significant advantages of using the Pattern format instead of the Element format to represent scalar values in the matrices when these values were boolean, such as matrices in the SNAP collection which represent graph datasets are boolean. In particular, Finch Pattern SpMV had for had an average speedup of X over TACO and X over the same Finch program, but with an Element format.
